---
title: "Report on Fuel Consumption of automobiles"
output: rmarkdown::tufte_handout
---

```{r loads.libraries, echo=FALSE, message=FALSE}
require('ggplot2')
require('gridExtra')
require('data.table')
require('ggthemes')
require('leaps')
require('xtable')
options(xtable.comment = FALSE)
theme.bartek <- theme_bw()+theme(text=element_text(size=6))

mtcars.dt <- data.table(mtcars)
mtcars.dt$am <- as.factor(mtcars.dt$am)
mtcars.dt$cyl <- as.factor(mtcars.dt$cyl)
mtcars.dt$gear <- as.factor(mtcars.dt$gear)
mtcars.dt$vs <- as.factor(mtcars.dt$vs)
mtcars.dt$carb <- as.factor(mtcars.dt$carb)
```

# Introduction

```{r boxplot1, fig.margin = TRUE, fig.cap = "Automatic vs. manual transmission.", echo=FALSE}
ggplot(mtcars.dt, aes(am, mpg)) + geom_boxplot() + theme_tufte() +
            scale_x_discrete(labels=c("manual", "automatic"))
```
```{r WTGimpact, fig.margin = TRUE, fig.cap = "Impact of weight, transmission and number of cylinders on fuel consumption.", echo=FALSE}
ggplot(mtcars.dt, aes(wt, mpg, colour=am)) + geom_point() + theme_tufte() +
    geom_text(aes(label=cyl, colour=NULL), vjust=-0.6) +
        geom_smooth(method='lm')
```

We investigate data that extracted from the 1974 _Motor Trend_ US
magazine. They comprises fuel consumption and 10 aspects of automobile
design and performance for 32 automobiles (1973-74 models). We are
particularly interested in the following questions:

* Is an automatic or manual transmission better for MPG?
* Quantify the MPG difference between automatic and manual
  transmissions.


The boxplot on Figure 1 suggests that manual transmission is better
for mpg. However, Figure 2 shows that actually weight or number of
gears can be a most important factor. Let's see.


# Model Selection

```{r r2statistic, fig.margin=TRUE, fig.cap="$R^2$ statistic for the best model with $n=1,\\ldots, 8$.", echo=FALSE}
regfit <- regsubsets(mpg~. ,mtcars.dt)
reg.summary <- summary(regfit)
ggplot(data=NULL, aes(x=1:length(reg.summary$rsq), y=reg.summary$rsq)) +
    geom_line() + theme_tufte() +
    labs(x='n', y="R2 statistic")
```


Let us try to identify the subset of the predictors that can be
related to the mpg response. For that we use `regsubsets` function
from `leaps` library which select the best model with \(n\)
predictors, for \(n=1 \ldots 8\). This way we can choose from quite
large amount of models. Additionally in Appendix we show plots of
few models with different sets of predictors.



```{r selection, results='asis', message=FALSE, echo=FALSE, warning=FALSE}
print(xtable(reg.summary$outmat), size="\\tiny")
```
If we take a look at Figure 3 where we plot $R^2$ statistic for the
best model with $n=1,..., 8$ predictors, we see that $n=5$ would be
the best choice. Therefore our regression model is the following has
predictors horsepower (hp), weight (wt), manual transmission (am=1),
six cylindes (cyl=6) and V/S (vs=1).
```{r weightVsAm, fig.margin=TRUE, fig.cap="Boxplots of weight", echo=FALSE}
ggplot(mtcars.dt, aes(x=am,y=wt)) + geom_boxplot() + theme_tufte()
```
```{r 5predictors, results='asis', echo=FALSE} 
fit <- lm(mpg~ hp + wt + am + I(cyl==6) + I(vs==1), data=mtcars.dt)
print(xtable(summary(fit)$coefficients, digits = 4), size="\\small")
```


# Conclusion
**Is an automatic or manual transmission better for MPG?**

```{r mpgVsWt, fig.margin=TRUE, fig.cap="Relation beteewn mpg and weight", echo=FALSE}
y <- mtcars$mpg
x <- mtcars$wt
fit <- lm(y~x)
b0 <- coef(fit)[1]
b1 <- coef(fit)[2]
fit.fun <- function(t) {b0+b1*t}
e <- y - fit.fun(x)
ggplot(data=NULL, aes(x, y)) + geom_point() +
    geom_smooth(method='lm') + theme_tufte()
```

```{r ResmpgVsWt, fig.margin=TRUE, fig.cap="Residual plot for regresion beteewn mpg and weight", echo=FALSE}
ggplot(data=NULL, aes(x,e)) + geom_point() +
    geom_hline(yintercept=0) +
        geom_pointrange(aes(ymin=pmin(e,0), ymax=pmax(e,0)))+
        theme_tufte()
```
In my opinion the results are not conclusive. This regression model
suggest that actually the most significant is characteristics are
horsepower and weight. One can clearly see that the more heavy cars
have automatic transmission (Figure 4). This could be an effect of
that economic car (in the sense of price) are smaller and tends to
have manual transmission while luxury cars are bigger and its
transmission is automatic.

**The MPG difference between automatic and manual transmissions**

```{r mpgVsHp, fig.margin=TRUE, fig.cap="Relation beteewn mpg and horsepower", echo=FALSE}
y <- mtcars$mpg
x <- mtcars$hp
fit <- lm(y~x)
b0 <- coef(fit)[1]
b1 <- coef(fit)[2]
fit.fun <- function(t) {b0+b1*t}
e <- y - fit.fun(x)
ggplot(data=NULL, aes(x, y)) + geom_point() +
    geom_smooth(method='lm') + theme_tufte()
```

```{r ResmpgVsHp, fig.margin=TRUE, fig.cap="Residual plot for regresion beteewn mpg and horsepower", echo=FALSE}
ggplot(data=NULL, aes(x,e)) + geom_point() +
    geom_hline(yintercept=0) +
        geom_pointrange(aes(ymin=pmin(e,0), ymax=pmax(e,0)))+
        theme_tufte()
```


The difference is clearly visible on Figure 1. Let us do however
_t-test_ to test if the difference in mean is significant.
```{r twoGroup, echo=FALSE}
g1 <- mtcars.dt[am==0,mpg]
g2 <- mtcars.dt[am==1,mpg]
t.test(mtcars[mtcars$am==0,1], mtcars[mtcars$am==1,1], paired=FALSE, var.equal=FALSE)
```

Hence the difference in mean is significant, however _correlation does
not imply causation_, and as we have mentioned above it is likely to
be caused by other factors like weight and horsepower. For example,
on Figure 5 we presented how mpg depends on weight and on Figure 6 the
residual plot for the linear regression. And on the next figures we
did the same for horsepower. These relations definitely make sense.

The code of this file is available on my github:

`https://github.com/sbartek/mpg_report`

\newpage

# Appendix

```{r, fig.width = 12, fig.height=8, fig.fullwidth = TRUE, fig.cap = "Pairs figure."}
pairs(mtcars.dt[,.(mpg,cyl, disp, hp, drat, wt, qsec, vs, am, gear, carb)], panel = panel.smooth)
```

```{r boxplot12, fig.height=2, fig.cap = "Fuel consumption. Automatic vs. manual transmission and impact of weight, transmission and number of cylinders.", fig.fullwidth = TRUE}
p1 <- ggplot(mtcars.dt, aes(am, mpg)) + geom_boxplot() +
    theme.bartek + scale_x_discrete(labels=c("manual", "automatic"))
p2 <- ggplot(mtcars.dt, aes(wt, mpg, colour=am)) + geom_point()+
    theme.bartek + theme(legend.position="none") +
    geom_text(aes(label=cyl, colour=NULL), vjust=-0.6, size=2) +
        geom_smooth(method='lm')
grid.arrange(p1, p2, ncol=2)
```


```{r, fig.cap = "Horsepower vs. mpg"}
ggplot(mtcars.dt, aes(hp, mpg, colour=am)) +
    geom_point() +
        geom_text(aes(label=cyl, colour=NULL), vjust=-0.6) +
        theme.bartek +
            geom_smooth(method='lm')
```

```{r, fig.cap="Displacement vs. mpg"}
ggplot(mtcars.dt, aes(disp, mpg, colour=am)) +
    geom_point() +
        theme.bartek +
            geom_smooth(method='lm')

```

[^git]: https://github.com/sbartek/mpg_report










